Active Learning for SAT-Solver Benchmarking

### 2022-09-27 paper/paper.tex
Add comments from PdF paper, part 2 (Sections 4-6)
... and slightly adapt text in some places (comments I directly resolved or
other stuff I noticed). Sections 1-3 are affected as well for consistency
reasons (e.g., add further line breaks between sentences and harmonize
usage of term "New-Solver Problem").

\begin{abstract}
  Experimentation is one of the main phases in algorithm engineering, including the propositional-sat\-is\-fia\-bi\-li\-ty problem (SAT).
  While the field of SAT-solver benchmarking is well established, traditional bench\-mark-set-selection approaches do not optimize for benchmark runtime.
  Benchmark-set selection is the process of choosing a subset of instances out of a pool to be representative and reliable in discriminating SAT solvers.
  This paper presents an active-learning-based benchmark-set-selec\-tion approach that solves the new-solver problem:
  Given a new solver, we want to find its ranking amidst its competitors with as little runtime as possible while still giving accurate ranking predictions.
  We can determine a new solver's rank regarding the SAT Competition 2022 Anniversary Track instances in about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset.
  We achieve this while still maintaining a prediction accuracy of about \SI{92}{\%} regarding the PAR-2 ranking of this solver if it were evaluated on the complete dataset.
  Additionally, we discuss the importance of instance families in the selection process. 
  Our tool provides the engineer of propositional-satisfiability solvers with a faster way of determining a new SAT-solver's performance in comparison to sampling the benchmark set by hand.

@@

\begin{definition}[New solver problem]
	Given a pool of instances $\mathcal{I}$ and solvers $\mathcal{A}$ with already known runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$ (timeout $\tau$ usually \SI{5000}{s}~\cite{FroleyksHIJS21}), the problem is to decide with as little resources (CPU time) as possible how a new solver $\hat{a} \notin \mathcal{A}$ performs in comparison to the existing ones by incrementally selecting a benchmark set $\tilde{\mathcal{I}}$ on which to evaluate $\hat{a}$.
\end{definition}

@@

To sum up our contribution, we take an existing benchmark set, i.e., the SAT Competition~2022 Anniversary Track~\cite{sat2022}, and predict a new solver's PAR2 score on this dataset with only a fraction of the runtime that would be needed to evaluate this solver on the complete set of instances.
%JB: I think we should expand a bit more on experimental design (how many instances/solvers, what do we compare), and the key results; I think it's okay if the intro becomes a bit longer (and in turn, the overview of the approach's features can be cut a bit)

@@

\section{Related Work}
This section discusses work in several related fields.
First, we take a look at traditional SAT-solver benchmarking.
Second, we explore how SAT-solver configuration systems deal with navigating possible configurations.
Third, we discuss an existing benchmark-set-selection approach that optimizes for benchmark runtime.
Finally, we show how active-learning techniques can help to tackle the aforementioned \textit{new-solver} problem.

@@

\begin{table}[tbp]
  \centering
  \caption{Support for desired features that are introduced in Sec.~\ref{sec:intro}. The shown approaches include traditional benchmark-set-selection techniques~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.}
  \label{tab:requirements}
  \vspace{0.2cm}
  \begin{tabular}{
    m{0.3\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
  }
    \hline
    Feature & Hoos \mbox{et. al.~\cite{HoosKSS13}} & SMAC~\cite{HutterHL11} & Matricon \mbox{et. al.~\cite{MatriconAFSH21}} & Our approach \\
    \hline
    Ranking & \cmark & \xmark & \cmark & \cmark \\
    Runtime optimization & \xmark & \cmark & \cmark & \cmark \\
    %Feature-based & \cmark & \cmark & \cmark & \cmark \\
    Incremental non-random sampling & \xmark & \xmark & \cmark & \cmark \\
    % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance sampling is done randomly.}
    Interpretability & \cmark & \xmark & \xmark & \cmark \\
    % \tablefootnote{While they provide a confidence measure, the strategy's output is still only binary, i.e., whether a solver A is better or worse than solver B.}
    Scalability & \cmark & \cmark & \xmark & \cmark \\
    \hline
  \end{tabular}
\end{table}

@@

\subsubsection{Active learning (AL).}
%JB: section does not make clear which of the following approaches we will use (reads more like fundamentals than related work)
The posed \textit{new solver problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \textit{new user problem}~\cite{RubensESK15}.
The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
%JB: no sure it is arises here as well; dilemma is typical for searching one optimal solution in continuous space; however, if we find an interesting/useful instance here, I'm not sure if we would continue picking similar instances ("exploting" instead of "exploring") just because we expect them to be useful as well

@@

\subsubsection{Summary.}
To sum up this section, we have shown both traditional and recent benchmark-set-selection approaches. Moreover, we have briefly outlined how active-learning techniques can help to incrementally select benchmark sets. In the following sections, we present an active-learning-based benchmark-set-selection approach that solves the \textit{new solver problem} while optimizing for runtime and being both interpretable and scalable.

@@

%\begin{table}[tbp]
%  \centering
%  \caption{PAR-2 ranking vs. discretized-runtime ranking for the SAT Competition~2022 Anniversary Track instances~\cite{sat2022} ($k = 3$; hierarchical clustering). Differences are highlighted with a gray background.}
%  \label{tab:discreteranking}
%  \begin{tabular}{
%    c
%    >{\centering\arraybackslash}p{1.8cm}p{3.6cm}
%    >{\centering\arraybackslash}p{1.8cm}p{3.6cm}
%  }
%    \hline
%    Rank & \multicolumn{2}{c}{PAR-2 Ranking} & \multicolumn{2}{c}{Cluster-Label Ranking} \\
%    \hline
%1   & 2806.400 & Kissat\_MAB\_ESA           & 1.172 & Kissat\_MAB\_ESA \\
%2   & 2810.959 & kissat-sc2022-bulky        & 1.183 & kissat-sc2022-bulky \\
%3   & 2830.143 & ekissat-mab-gb-db          & 1.185 & ekissat-mab-gb-db \\
%\cellcolor{gray!15} 4   & \cellcolor{gray!15} 2833.081 & \cellcolor{gray!15} Kissat\_MAB\_UCB           & \cellcolor{gray!15} 1.186 & \cellcolor{gray!15} kissat\_inc \\
%\cellcolor{gray!15} 5   & \cellcolor{gray!15} 2833.397 & \cellcolor{gray!15} kissat\_inc                & \cellcolor{gray!15} 1.186 & \cellcolor{gray!15} Kissat\_MAB\_UCB \\
%6   & 2841.734 & ekissat-mab-db-v1          & 1.192 & ekissat-mab-db-v1 \\
%7   & 2842.666 & Kissat\_MAB\_MOSS          & 1.192 & Kissat\_MAB\_MOSS \\
%8   & 2857.219 & Kissat\_MAB-HyWalk         & 1.195 & Kissat\_MAB-HyWalk \\
%9   & 2866.960 & kissat-sc2022-light        & 1.120 & kissat-sc2022-light \\
%10  & 2896.371 & kissat-els-v2              & 1.216 & kissat-els-v2 \\
%11  & 2950.179 & hKis-unsat                 & 1.229 & hKis-unsat \\
%12  & 2964.729 & Kissat\_adaptive\_restart  & 1.234 & Kissat\_adaptive\_restart \\
%13  & 2974.990 & SeqFROST-NoExtend          & 1.247 & SeqFROST-NoExtend \\
%14  & 3010.642 & Cadical\_ESA               & 1.248 & Cadical\_ESA \\
%\cellcolor{gray!15} 15  & \cellcolor{gray!15} 3012.152 & \cellcolor{gray!15} kissat-els-v1              & \cellcolor{gray!15} 1.258 & \cellcolor{gray!15} CadicalReorder \\
%\cellcolor{gray!15} 16  & \cellcolor{gray!15} 3028.922 & \cellcolor{gray!15} CadicalReorder             & \cellcolor{gray!15} 1.264 & \cellcolor{gray!15} kissat-els-v1 \\
%\cellcolor{gray!15} 17  & \cellcolor{gray!15} 3042.540 & \cellcolor{gray!15} cadical\_rel\_Scavel       & \cellcolor{gray!15} 1.278 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V1 \\
%\cellcolor{gray!15} 18  & \cellcolor{gray!15} 3082.293 & \cellcolor{gray!15} kissat\_relaxed            & \cellcolor{gray!15} 1.283 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V2 \\
%\cellcolor{gray!15} 19  & \cellcolor{gray!15} 3089.304 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V1          & \cellcolor{gray!15} 1.298 & \cellcolor{gray!15} kissat\_relaxed \\
%\cellcolor{gray!15} 20  & \cellcolor{gray!15} 3095.118 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V2          & \cellcolor{gray!15} 1.361 & \cellcolor{gray!15} cadical\_rel\_Scavel \\
%21  & 3272.674 & glucose-reboot             & 1.379 & glucose-reboot \\
%\cellcolor{gray!15} 22  & \cellcolor{gray!15} 3287.322 & \cellcolor{gray!15} LStech-Maple-HyWalk        & \cellcolor{gray!15} 1.414 & \cellcolor{gray!15} hCaD\_V1-psids \\
%23  & 3288.798 & LSTech\_Maple              & 1.446 & LSTech\_Maple \\
%\cellcolor{gray!15} 24  & \cellcolor{gray!15} 3399.281 & \cellcolor{gray!15} SLIME-SC-2022-beta         & \cellcolor{gray!15} 1.469 & \cellcolor{gray!15} LStech-Maple-HyWalk \\
%\cellcolor{gray!15} 25  & \cellcolor{gray!15} 3410.345 & \cellcolor{gray!15} SLIME-SC-2022              & \cellcolor{gray!15} 1.470 & \cellcolor{gray!15} SLIME-SC-2022-beta \\
%\cellcolor{gray!15} 26  & \cellcolor{gray!15} 3430.403 & \cellcolor{gray!15} hCaD\_V1-psids             & \cellcolor{gray!15} 1.523 & \cellcolor{gray!15} SLIME-SC-2022 \\
%27  & 3504.244 & MapleLCMDistChrBt-DL-v3    & 1.542 & MapleLCMDistChrBt-DL-v3 \\
%28  & 4750.808 & IsaSAT                     & 2.062 & IsaSAT \\
%    \hline
%  \end{tabular}
%\end{table}

@@

\subsection{Optimization Goal}
\label{sec:goal}
As already stated in the introductory section, this work addresses the \textit{new solver problem}.
By running Alg.~\ref{algALBenchmark}, we obtain a prediction model $\mathcal{M}$ that provides us with a scoring function \textit{score}.
Also, it outputs the fraction of runtime that is needed to arrive at its conclusion.
First and foremost, our goal is to provide the engineer of new SAT solvers with an accurate ranking.
We define the ranking accuracy $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking.
For now, we exclude the equality of solvers since data shows that the PAR-2 scores of all our solvers are different.
So, possible ranking decisions may be, for example, solver $a$ is better than $b$ or $b$ is better than $a$.
The ranking accuracy is affected by whether this decision is correctly made.
Matricon et. al.~\cite{MatriconAFSH21} also evaluate their approach by the fraction of solver pairs that is correctly decided.
Second, we also have to optimize for runtime.
The fraction of runtime that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
In that sense, we want to find an approach that maximizes
\begin{equation}
  O_\delta := \delta O_{\operatorname{acc}} + \left(1 - \delta\right) \left(1 - O_{\operatorname{rt}}\right) \enspace \textrm{,}
  \label{eq:opt}
\end{equation} 
whereby $\delta \in \left[0, 1\right]$ allows for linear weighting between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on a $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.

@@

    $O_{\operatorname{rt}} \leftarrow \operatorname{computeRuntimeFractionUsed}\!\left( runtime,\, \mathcal{I},\, \mathcal{S},\, r \right)$

@@

\subsection{Data}
In our experiments, we work with the SAT~Competition~2022 Anniversary Track instances~\cite{sat2022}.
The dataset consists of 5355 instances with respective runtime data.
It has complete runtimes of 28 solvers.
We use the GBD metadata database~\cite{IserS18} to provide us with the aforementioned instance features, problem instances, and solver runtimes.
Thereby, all our approaches make use of the 56 base features that are maintained in the \textit{base} database of GBD~\cite{IserS18}.
They are inspired by the SATzilla features~\cite{features}.
Those features comprise general instance-size features and graph-representation features among others.
All features are numeric and fortunately free of any missing values.
We drop 10 out of 56 features because of zero variance.
For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance's family.
All instance families that are too \textit{small}, i.e., \SI{10}{\%} of it is less than one, are put into one meta-family for stratification.
This smaller dataset allows for a more extensive exploration of hyper-parameter space.

\subsection{Hyper-parameters}
\label{sec:hyper}
Given Alg.~\ref{algALBenchmark}, there are several possible instantiations for the three phases, i.e., \textit{selection}, \textit{stopping}, and \textit{ranking}.
Also, there are different choices for the runtime-label prediction model.
Note that we are not considering all previously listed approaches since a grid search of all combinations would be infeasible. Rather, we filter approaches based on preliminary experimental results (cf. Sec.~\ref{sec:evalprel}) and do the main end-to-end experiments only with a subset.
The end-to-end experiment configurations are given below.

\subsubsection{Ranking.}
Regarding \textit{ranking} (cf. Sec.~\ref{sec:main1}), we experiment with the following approaches, including our used hyper-parameter values:
\begin{itemize}
  \item Observed PAR-2 ranking of already sampled instances
  \item Predicted ranking induced by runtime-label predictions
  \begin{itemize}
    \item History size: consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability.
    \item Fallback threshold: if the difference of ranking scores drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
  \end{itemize}
\end{itemize}

\subsubsection{Selection.}
For \textit{selection} (cf. Sec.~\ref{sec:sampling1}), we experiment with the following methods, including our used hyper-parameter values:
\begin{itemize}
  \item Random sampling 
  \item Model-based uncertainty sampling
  \begin{itemize}
    \item Fallback threshold: use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
    \item Runtime scaling: prefer instances with the greatest uncertainty per average runtime (\texttt{True} or \texttt{False}).
  \end{itemize}

@@

\subsubsection{Stopping.}
For \textit{stopping} decisions (cf. Sec.~\ref{sec:main3}), we experiment with the following criteria, including our used hyper-parameter values:
\begin{itemize}
  \item Fixed subset size of \SI{10}{\%} or \SI{20}{\%} of instances
  \item Ranking convergence criterion
  \begin{itemize}
    \item Minimum amount: sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
    \item Convergence duration: stop if the predicted ranking stays the same for the last \SI{1}{\%} or \SI{2}{\%} of sampled instances.
  \end{itemize}

@@

%the proposed runtime label ranking, (2) the PAR-2 subsample ranking, and (3) the runtime label ranking that uses the PAR-2 subsample ranking as a fallback when performance differences drop below a threshold. Possible values are \SI{0.01}{}, \SI{0.05}{}, and \SI{0.1}{}.
%Also, we experiment with different lengths of the prediction history that is used for stabilizing the ranking. Possible values are 10, 20, 30, 40, and 50.

\subsection{Implementation Details}
For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}}.
It is implemented in \textsc{Python} using \textit{scikit-learn}~\cite{scikit-learn} for making predictions and \textit{gbd-tools}~\cite{IserS18} for SAT-instance retrieval.

>>>>

\begin{abstract}
  Experimentation is one of the main phases in algorithm engineering, including the propositional-sat\-is\-fia\-bi\-li\-ty problem (SAT).
  While the field of SAT-solver benchmarking is well established, traditional bench\-mark-set-selection approaches do not optimize for benchmark runtime.
  Benchmark-set selection is the process of choosing a subset of instances out of a pool to be representative and reliable in discriminating SAT solvers.
  This paper presents an active-learning-based benchmark-set-selec\-tion approach that solves the \textit{New-Solver Problem}:
  Given a new solver, we want to find its ranking amidst its competitors with as little runtime as possible while still giving accurate ranking predictions.
  We can determine a new solver's rank regarding the SAT Competition 2022 Anniversary Track instances in about \SI{10}{\%} of the time it would take to run the solver on all instances of this dataset.
  We achieve this while still maintaining a prediction accuracy of about \SI{92}{\%} regarding the PAR-2 ranking of this solver if it were evaluated on the complete dataset.
  Additionally, we discuss the importance of instance families in the selection process. 
  Our tool provides the engineer of propositional-satisfiability solvers with a faster way of determining a new SAT-solver's performance in comparison to sampling the benchmark set by hand.

@@

\begin{definition}[New-Solver Problem]
	Given a pool of instances $\mathcal{I}$ and solvers $\mathcal{A}$ with already known runtimes $r\!: \mathcal{A} \times \mathcal{I} \rightarrow \left[0, \tau\right] \cup \left\lbrace \perp \right\rbrace$ (timeout $\tau$ usually \SI{5000}{s}~\cite{FroleyksHIJS21}), the problem is to decide with as little resources (CPU time) as possible how a new solver $\hat{a} \notin \mathcal{A}$ performs in comparison to the existing ones by incrementally selecting a benchmark set $\tilde{\mathcal{I}}$ on which to evaluate $\hat{a}$.
	\label{def:new-solver-problem}
\end{definition}

@@

To sum up our contribution, we take an existing benchmark set, i.e., the SAT Competition~2022 Anniversary Track~\cite{sat2022}, and predict a new solver's PAR-2 score on this dataset with only a fraction of the runtime that would be needed to evaluate this solver on the complete set of instances.
%JB: I think we should expand a bit more on experimental design (how many instances/solvers, what do we compare), and the key results; I think it's okay if the intro becomes a bit longer (and in turn, the overview of the approach's features can be cut a bit)

@@

\section{Related Work}
This section discusses work in several related fields.
First, we take a look at traditional SAT-solver benchmarking.
Second, we explore how SAT-solver configuration systems deal with navigating possible configurations.
Third, we discuss an existing benchmark-set-selection approach that optimizes for benchmark runtime.
Finally, we show how active-learning techniques can help to tackle the aforementioned \textit{New-Solver Problem}.

@@

\begin{table}[tbp]
  \centering
  \caption{
  	Support for desired features that are introduced in Sec.~\ref{sec:intro}.
  	The shown approaches include traditional benchmark-set-selection techniques~\cite{HoosKSS13}, an algorithm configuration system~\cite{HutterHL11}, an existing active-learning approach~\cite{MatriconAFSH21}, as well as our approach.
  }
  \label{tab:requirements}
  \vspace{0.2cm}
  \begin{tabular}{
    m{0.3\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
    >{\centering\arraybackslash}m{0.15\textwidth}
  }
    \hline
    Feature & Hoos \mbox{et. al.~\cite{HoosKSS13}} & SMAC~\cite{HutterHL11} & Matricon \mbox{et. al.~\cite{MatriconAFSH21}} & Our approach \\
    \hline
    Ranking & \cmark & \xmark & \cmark & \cmark \\
    Runtime optimization & \xmark & \cmark & \cmark & \cmark \\
    %Feature-based & \cmark & \cmark & \cmark & \cmark \\
    Incremental non-random sampling & \xmark & \xmark & \cmark & \cmark \\
    % \tablefootnote{While using a model-based selection strategy to select the next configuration to evaluate, the instance sampling is done randomly.}
    Interpretability & \cmark & \xmark & \xmark & \cmark \\
    % \tablefootnote{While they provide a confidence measure, the strategy's output is still only binary, i.e., whether a solver A is better or worse than solver B.}
    Scalability & \cmark & \cmark & \xmark & \cmark \\
    \hline
  \end{tabular}
\end{table}

@@

\subsubsection{Active learning (AL).}
%JB: section does not make clear which of the following approaches we will use (reads more like fundamentals than related work)
The posed \textit{New-Solver Problem} has stark similarities to the well-studied field of active learning within recommender systems, especially the \textit{new user problem}~\cite{RubensESK15}.
The classic exploration-exploitation dilemma arises since, on the one hand, we want to maximize the utility an instance provides to our model and, on the other hand, minimize the cost (CPU time) that is associated with its acquisition.
%JB: no sure it is arises here as well; dilemma is typical for searching one optimal solution in continuous space; however, if we find an interesting/useful instance here, I'm not sure if we would continue picking similar instances ("exploting" instead of "exploring") just because we expect them to be useful as well

@@

\subsubsection{Summary.}
To sum up this section, we have shown both traditional and recent benchmark-set-selection approaches.
Moreover, we have briefly outlined how active-learning techniques can help to incrementally select benchmark sets.
In the following sections, we present an active-learning-based benchmark-set-selection approach that solves the \textit{New-Solver Problem} while optimizing for runtime and being both interpretable and scalable.

@@

%\begin{table}[tbp]
%  \centering
%   \caption{
%     PAR-2 ranking vs. discretized-runtime ranking for the SAT Competition~2022 Anniversary Track instances~\cite{sat2022} ($k = 3$; hierarchical clustering).
%	  Differences are highlighted with a gray background.
%	}
%  \label{tab:discreteranking}
%  \begin{tabular}{
%    c
%    >{\centering\arraybackslash}p{1.8cm}p{3.6cm}
%    >{\centering\arraybackslash}p{1.8cm}p{3.6cm}
%  }
%    \hline
%    Rank & \multicolumn{2}{c}{PAR-2 Ranking} & \multicolumn{2}{c}{Cluster-Label Ranking} \\
%    \hline
%1   & 2806.400 & Kissat\_MAB\_ESA           & 1.172 & Kissat\_MAB\_ESA \\
%2   & 2810.959 & kissat-sc2022-bulky        & 1.183 & kissat-sc2022-bulky \\
%3   & 2830.143 & ekissat-mab-gb-db          & 1.185 & ekissat-mab-gb-db \\
%\cellcolor{gray!15} 4   & \cellcolor{gray!15} 2833.081 & \cellcolor{gray!15} Kissat\_MAB\_UCB           & \cellcolor{gray!15} 1.186 & \cellcolor{gray!15} kissat\_inc \\
%\cellcolor{gray!15} 5   & \cellcolor{gray!15} 2833.397 & \cellcolor{gray!15} kissat\_inc                & \cellcolor{gray!15} 1.186 & \cellcolor{gray!15} Kissat\_MAB\_UCB \\
%6   & 2841.734 & ekissat-mab-db-v1          & 1.192 & ekissat-mab-db-v1 \\
%7   & 2842.666 & Kissat\_MAB\_MOSS          & 1.192 & Kissat\_MAB\_MOSS \\
%8   & 2857.219 & Kissat\_MAB-HyWalk         & 1.195 & Kissat\_MAB-HyWalk \\
%9   & 2866.960 & kissat-sc2022-light        & 1.120 & kissat-sc2022-light \\
%10  & 2896.371 & kissat-els-v2              & 1.216 & kissat-els-v2 \\
%11  & 2950.179 & hKis-unsat                 & 1.229 & hKis-unsat \\
%12  & 2964.729 & Kissat\_adaptive\_restart  & 1.234 & Kissat\_adaptive\_restart \\
%13  & 2974.990 & SeqFROST-NoExtend          & 1.247 & SeqFROST-NoExtend \\
%14  & 3010.642 & Cadical\_ESA               & 1.248 & Cadical\_ESA \\
%\cellcolor{gray!15} 15  & \cellcolor{gray!15} 3012.152 & \cellcolor{gray!15} kissat-els-v1              & \cellcolor{gray!15} 1.258 & \cellcolor{gray!15} CadicalReorder \\
%\cellcolor{gray!15} 16  & \cellcolor{gray!15} 3028.922 & \cellcolor{gray!15} CadicalReorder             & \cellcolor{gray!15} 1.264 & \cellcolor{gray!15} kissat-els-v1 \\
%\cellcolor{gray!15} 17  & \cellcolor{gray!15} 3042.540 & \cellcolor{gray!15} cadical\_rel\_Scavel       & \cellcolor{gray!15} 1.278 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V1 \\
%\cellcolor{gray!15} 18  & \cellcolor{gray!15} 3082.293 & \cellcolor{gray!15} kissat\_relaxed            & \cellcolor{gray!15} 1.283 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V2 \\
%\cellcolor{gray!15} 19  & \cellcolor{gray!15} 3089.304 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V1          & \cellcolor{gray!15} 1.298 & \cellcolor{gray!15} kissat\_relaxed \\
%\cellcolor{gray!15} 20  & \cellcolor{gray!15} 3095.118 & \cellcolor{gray!15} CaDiCaL\_DVDL\_V2          & \cellcolor{gray!15} 1.361 & \cellcolor{gray!15} cadical\_rel\_Scavel \\
%21  & 3272.674 & glucose-reboot             & 1.379 & glucose-reboot \\
%\cellcolor{gray!15} 22  & \cellcolor{gray!15} 3287.322 & \cellcolor{gray!15} LStech-Maple-HyWalk        & \cellcolor{gray!15} 1.414 & \cellcolor{gray!15} hCaD\_V1-psids \\
%23  & 3288.798 & LSTech\_Maple              & 1.446 & LSTech\_Maple \\
%\cellcolor{gray!15} 24  & \cellcolor{gray!15} 3399.281 & \cellcolor{gray!15} SLIME-SC-2022-beta         & \cellcolor{gray!15} 1.469 & \cellcolor{gray!15} LStech-Maple-HyWalk \\
%\cellcolor{gray!15} 25  & \cellcolor{gray!15} 3410.345 & \cellcolor{gray!15} SLIME-SC-2022              & \cellcolor{gray!15} 1.470 & \cellcolor{gray!15} SLIME-SC-2022-beta \\
%\cellcolor{gray!15} 26  & \cellcolor{gray!15} 3430.403 & \cellcolor{gray!15} hCaD\_V1-psids             & \cellcolor{gray!15} 1.523 & \cellcolor{gray!15} SLIME-SC-2022 \\
%27  & 3504.244 & MapleLCMDistChrBt-DL-v3    & 1.542 & MapleLCMDistChrBt-DL-v3 \\
%28  & 4750.808 & IsaSAT                     & 2.062 & IsaSAT \\
%    \hline
%  \end{tabular}
%\end{table}

@@

\subsection{Optimization Goal}
\label{sec:goal}
As already stated in the introductory section, this work addresses the \textit{New-Solver Problem} (cf.~Def.~\ref{def:new-solver-problem}).
By running Alg.~\ref{algALBenchmark}, we obtain a prediction model $\mathcal{M}$ that provides us with a scoring function \textit{score}.
Also, it outputs the fraction of runtime that is needed to arrive at its conclusion.
First and foremost, our goal is to provide the engineer of new SAT solvers with an accurate ranking.
We define the \textit{ranking accuracy} $O_{\operatorname{acc}} \in \left[0, 1\right]$ (higher is better) by the fraction of pairs $\left(\hat{a}, a\right)$ for all $a \in \mathcal{A}$ that are decided correctly by the given ranking.
For now, we exclude the equality of solvers since data shows that the PAR-2 scores of all our solvers are different.
So, possible ranking decisions may be, for example, solver $a$ is better than $b$ or $b$ is better than $a$.
The ranking accuracy is affected by whether this decision is correctly made.
Matricon et. al.~\cite{MatriconAFSH21} also evaluate their approach by the fraction of solver pairs that is correctly decided.
Second, we also have to optimize for runtime.
The \textit{fraction of runtime} that the algorithm needs to arrive at its conclusion is denoted by $O_{\operatorname{rt}} \in \left[0, 1\right]$ (lower is better).
This metric puts the runtime summed over the sampled instances in relation to the runtime summed over all instances in the dataset.
Overall, we want to find an approach that maximizes
\begin{equation}
  O_\delta := \delta O_{\operatorname{acc}} + \left(1 - \delta\right) \left(1 - O_{\operatorname{rt}}\right) \enspace \textrm{,}
  \label{eq:opt}
\end{equation} 
whereby $\delta \in \left[0, 1\right]$ allows for linear weighting between the two optimization goals $O_{\operatorname{acc}}$ and $O_{\operatorname{rt}}$.
Plotting the approaches that maximize $O_\delta$ for all $\delta \in \left[0, 1\right]$ on a $O_{\operatorname{rt}}$-$O_{\operatorname{acc}}$-diagram provides us with a Pareto front of the best approaches for different optimization-goal weightings.

@@

    $O_{\operatorname{rt}} \leftarrow \operatorname{computeRuntimeFractionUsed}\!\left( runtime,\, \mathcal{I},\, \mathcal{S},\, r \right)$
    %JB: do we need to call a function here or can't we directly write how this is computed (define a variable starting at 0 to which the runtime fraction is added each iteration)

@@

\subsection{Data}
In our experiments, we work with the SAT~Competition~2022 Anniversary Track instances~\cite{sat2022}.
The dataset consists of 5355 instances with respective runtime data.
It has complete runtimes of 28 solvers.
We use the GBD metadata database~\cite{IserS18} to provide us with the aforementioned instance features, problem instances, and solver runtimes.
Thereby, all our approaches make use of the 56 base features that are maintained in the \textit{base} database of GBD~\cite{IserS18}.
They are inspired by the SATzilla features~\cite{features}.
Those features comprise general instance-size features and graph-representation features among others.
All features are numeric and fortunately free of any missing values.
We drop 10 out of 56 features because of zero variance.
For hyper-parameter tuning, we randomly sample \SI{10}{\%} of the complete set of 5355 instances with stratification regarding the instance's family.
All instance families that are too \textit{small}, i.e., \SI{10}{\%} of them corresponds to less than one instance, are put into one meta-family for stratification.
This smaller dataset allows for a more extensive exploration of hyper-parameter space.

\subsection{Hyper-parameters}
\label{sec:hyper}
Given Alg.~\ref{algALBenchmark}, there are several possible instantiations for the three phases, i.e., \textit{selection}, \textit{stopping}, and \textit{ranking}.
Also, there are different choices for the runtime-label prediction model.
Note that we are not considering all previously listed approaches since a grid search of all combinations would be infeasible.
Rather, we filter approaches based on preliminary experimental results (cf. Sec.~\ref{sec:evalprel}) and do the main end-to-end experiments only with a subset.
The end-to-end experiment configurations are given below.
%JB: still interesting to see which approaches were considered in preliminary experiments; could also list them here and indicate somehow which methods were only used preliminarily (e.g., by putting "(preliminray experiments only)" or some shorter label/sign behind them)

\subsubsection{Ranking.}
Regarding \textit{ranking} (cf. Sec.~\ref{sec:main1}), we experiment with the following approaches, including our used hyper-parameter values:
%JB: as these approaches are mostly new (Section 3.1 only explains prediction-based ranking in general, neither observed ranking nor history nor fallback), I'm not sure if we should provide further motivation/explanation here); also applies to other components of experimental design
\begin{itemize}
  \item Observed PAR-2 ranking of already sampled instances
  \item Predicted ranking induced by runtime-label predictions
  \begin{itemize}
    \item History size: consider the latest 1, 10, 20, 30, or 40 predictions within a voting approach for stability.
    \item Fallback threshold: if the difference of ranking scores drops below \SI{0.01}, \SI{0.05}, or \SI{0.1}, use the partially observed PAR-2 ranking as a tie-breaker.
    %JB: "difference of ranking" between a pair of new solver and another solver? or has this phrase a different meaning?
  \end{itemize}
\end{itemize}

\subsubsection{Selection.}
For \textit{selection} (cf. Sec.~\ref{sec:sampling1}), we experiment with the following methods, including our used hyper-parameter values:
%JB: sampling section describes more methods than the ones listed below, I guess some got dropped after preliminary experiments
\begin{itemize}
  \item Random sampling 
  \item Model-based uncertainty sampling
  \begin{itemize}
    \item Fallback threshold: use random sampling for the first \SI{0}{\%}, \SI{5}{\%}, \SI{10}{\%}, \SI{15}{\%}, or \SI{20}{\%} of instances to explore the instance space.
    \item Runtime scaling: prefer instances with the greatest uncertainty per average runtime (\texttt{True} or \texttt{False}).
    %JB: how does this scaling work exactly? divide runtimes by average runtime of all solvers on that instance?
  \end{itemize}

@@

\subsubsection{Stopping.}
For \textit{stopping} decisions (cf. Sec.~\ref{sec:main3}), we experiment with the following criteria, including our used hyper-parameter values:
\begin{itemize}
  \item Fixed subset size of \SI{10}{\%} or \SI{20}{\%} of instances
  \item Ranking convergence criterion
  \begin{itemize}
    \item Minimum amount: sample at least \SI{2}{\%}, \SI{8}{\%}, \SI{10}{\%}, or \SI{12}{\%} of instances before applying the criterion.
    \item Convergence duration: stop if the predicted ranking stays the same for the last \SI{1}{\%} or \SI{2}{\%} of sampled instances.
    %JB: why this fraction and not an absolute number of instances? the definition here might make stopping in early iterations rather likely, as just very few consecutive instances are needed (which might happen by coincidence; e.g., for ~5k instances, 2% minimum amount-> ~100 instances, 1% convergence duration -> ~1 instance), while the criterion becomes harder in later iterations
  \end{itemize}

@@

%the proposed runtime label ranking, (2) the PAR-2 subsample ranking, and (3) the runtime label ranking that uses the PAR-2 subsample ranking as a fallback when performance differences drop below a threshold.
%Possible values are \SI{0.01}{}, \SI{0.05}{}, and \SI{0.1}{}.
%Also, we experiment with different lengths of the prediction history that is used for stabilizing the ranking.
%Possible values are 10, 20, 30, 40, and 50.

\subsection{Implementation Details}
For reproducibility, our source code, all experiments, and data are available on GitHub\footnote{\url{https://github.com/mathefuchs/al-for-sat-solver-benchmarking}}.
Our code is implemented in \textsc{Python} using \textit{scikit-learn}~\cite{scikit-learn} for making predictions and \textit{gbd-tools}~\cite{IserS18} for SAT-instance retrieval.
